Implicit yaddayadda

### 2011-10-26 papers/PLDI2012/implicitization.tex
and -> virgule
Anglais: actual -> current, et inside of -> inside
"inside of" existe aussi si inside est un nom (the inside of blabla),
mais sauf erreur de ma part, il était de trop là où je l'ai enlevé.

Most SMT solvers, including Z3%
\footnote{\url{http://research.microsoft.com/en-us/um/redmond/projects/z3/}}
and Yices,%
\footnote{\url{http://yices.csl.sri.com/}}
and all those supporting the full SMTLIB2 standard \cite{BarST-SMTLIB},
offer an \emph{incremental} interface: the client program specifies the formula as an initially empty conjunction, to which additional constraints are added, and calls a ``check'' function answering whether it is satisfiable; 
it may then backtrack some of the constraints and add other ones without restarting from scratch.

@@

More generally, this analysis method considers an ascending sequence of subsets of the transitions in the loop body (left side of Fig.~\ref{fig:multigraph});
for each subset, an inductive invariant is computed for the program restricted to it.
The starting subset are the transitions reachable in one step from the loop initialization.
If for a given subset $S$ in the sequence, no transitions outside of $S$ are reachable from the inductive invariant attached to $S$, then iterations stop;
otherwise, add these transitions to $S$ and iterate more.
Termination ensues from the finiteness of the control-flow graph.

@@

In order to find focus paths, we solve an SMT formula which is satisfiable when
there exist a path starting at a point $p_i \in P_R$ in a state included in the
actual invariant candidate $X_i$, and arriving at a point $p_j \in P_R$ in a
state outside of $X_j$. In this case, we construct this path using the model and
update $X_j$. When $p_i = p_j$, meaning that the path is actually a self-loop,
we can apply a widening/narrowing sequence, or even compute the transitive
closure of the loop using abstract acceleration. % TODO citation laure gonnord

@@

In order to find which path to focus on, we construct an SMT formula $f(p_i)$, whose
model when satisfiable is a path that starts in $p_i$, goes to a successor $p_j
\in P_R$ of $p_i$, such that the image of $X_{i}^s$ by the path transformation
is not included in the actual $X_{j}^s$.
Intuitively, such a path makes the abstract value $X_{j}^s$ grow, and thus is
an interesting path to focus on. We loop until the formula becomes unsatisfiable,
meaning that the analysis of $p_i$ is finished.

@@

\begin{enumerate}
	\item The first subset of paths is depicted on Figure
		\ref{fig:example-graph} Step 1. We apply Path-Focusing on this graph,
		and discover for $p_5$ the inductive invariant $\lstinline|x_old| = 0$.
	\item
		We then compute the set $P'$ of paths that have to be added into the
		next subgraph. 
		The image of $\lstinline|x_old| = 0$ by the path that goes from $p_5$ to
		itself, and that goes through the \emph{else} branch of each 
		\emph{if-then-else}, is $-10 \leq \lstinline|x_old| \leq 10$. This path
		is then added to our subgraph. Moreover, there is no other path whose
		image is not in $-10 \leq \lstinline|x_old| \leq 10$.
		Additionally, the path that goes from $p_5$ to
		$p_11$, that goes through the \emph{else} branch of each 
		\emph{if-then-else} at lines 7,8 and 9, is also 
		$-10 \leq \lstinline|x_old| \leq 10$, and thus is also added into the
		new subgraph. Again, there is no new path that goes to $p_11$ with an
		image outside of $-10 \leq \lstinline|x_old| \leq 10$.
	\item
		We apply Path Focusing on this new subgraph (\ref{fig:example-graph}
		Step 2). When we have reached an inductive invariant, we narrow, and the
		result for both $p_5$ and $p_11$ is the polyhedron 
		$-10000 \leq \lstinline|x_old| \leq 10000$.
	\item Finally, we compute the new subgraph. The SMT-solver does not find any
		new path that makes the abstract values grow, and the algorithm
		terminates.
\end{enumerate}

>>>>

Most SMT solvers, including Z3,%
\footnote{\url{http://research.microsoft.com/en-us/um/redmond/projects/z3/}}
Yices,%
\footnote{\url{http://yices.csl.sri.com/}}
and all those supporting the full SMTLIB2 standard \cite{BarST-SMTLIB},
offer an \emph{incremental} interface: the client program specifies the formula as an initially empty conjunction, to which additional constraints are added, and calls a ``check'' function answering whether it is satisfiable; 
it may then backtrack some of the constraints and add other ones without restarting from scratch.

@@

More generally, this analysis method considers an ascending sequence of subsets of the transitions in the loop body (left side of Fig.~\ref{fig:multigraph});
for each subset, an inductive invariant is computed for the program restricted to it.
The starting subset are the transitions reachable in one step from the loop initialization.
If for a given subset $S$ in the sequence, no transitions outside $S$ are reachable from the inductive invariant attached to $S$, then iterations stop;
otherwise, add these transitions to $S$ and iterate more.
Termination ensues from the finiteness of the control-flow graph.

@@

In order to find focus paths, we solve an SMT formula which is satisfiable when
there exist a path starting at a point $p_i \in P_R$ in a state included in the
current invariant candidate $X_i$, and arriving at a point $p_j \in P_R$ in a
state outside $X_j$. In this case, we construct this path using the model and
update $X_j$. When $p_i = p_j$, meaning that the path is actually a self-loop,
we can apply a widening/narrowing sequence, or even compute the transitive
closure of the loop using abstract acceleration. % TODO citation laure gonnord

@@

In order to find which path to focus on, we construct an SMT formula $f(p_i)$, whose
model when satisfiable is a path that starts in $p_i$, goes to a successor $p_j
\in P_R$ of $p_i$, such that the image of $X_{i}^s$ by the path transformation
is not included in the current $X_{j}^s$.
Intuitively, such a path makes the abstract value $X_{j}^s$ grow, and thus is
an interesting path to focus on. We loop until the formula becomes unsatisfiable,
meaning that the analysis of $p_i$ is finished.

@@

\begin{enumerate}
	\item The first subset of paths is depicted on Figure
		\ref{fig:example-graph} Step 1. We apply Path-Focusing on this graph,
		and discover for $p_5$ the inductive invariant $\lstinline|x_old| = 0$.
	\item
		We then compute the set $P'$ of paths that have to be added into the
		next subgraph. 
		The image of $\lstinline|x_old| = 0$ by the path that goes from $p_5$ to
		itself, and that goes through the \emph{else} branch of each 
		\emph{if-then-else}, is $-10 \leq \lstinline|x_old| \leq 10$. This path
		is then added to our subgraph. Moreover, there is no other path whose
		image is not in $-10 \leq \lstinline|x_old| \leq 10$.
		Additionally, the path that goes from $p_5$ to
		$p_11$, that goes through the \emph{else} branch of each 
		\emph{if-then-else} at lines 7,8 and 9, is also 
		$-10 \leq \lstinline|x_old| \leq 10$, and thus is also added into the
		new subgraph. Again, there is no new path that goes to $p_11$ with an
		image outside $-10 \leq \lstinline|x_old| \leq 10$.
	\item
		We apply Path Focusing on this new subgraph (\ref{fig:example-graph}
		Step 2). When we have reached an inductive invariant, we narrow, and the
		result for both $p_5$ and $p_11$ is the polyhedron 
		$-10000 \leq \lstinline|x_old| \leq 10000$.
	\item Finally, we compute the new subgraph. The SMT-solver does not find any
		new path that makes the abstract values grow, and the algorithm
		terminates.
\end{enumerate}
